Computing Exact Restricted Isometry Constants via Mixed Integer Semidefinite Programming

### 2016-03-22 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
several minor things, handled marked paragraphs, added framework paper

\noindent
In this section we demonstrate the applicability of the proposed MISDP-formulation and use it to
assess the quality of the relaxations proposed in~\cite{Asp08} and~\cite{Asp07}.
For solving the mixed integer semidefinite programs, we use the MISDP solver SCIP-SDP~\cite{SCIP-SDP},
originally developed in~\cite{Mar13} and~\cite{MS12}. It combines the
branch and bound framework SCIP 3.2.1 \cite{SCIP} with 
interior-point SDP-solvers, in our case SDPA~7.3.8~\cite{SDPA6,SDPA7}.

@@

We use a testset consisting of 63 matrices of the following seven
different types: band matrices,
with band size three or five with entries within the band chosen uniformly in $\{0,1\}$, binary matrices with all entries chosen 
uniformly in $\{0,1\}$, Gaussian-distributed matrices, and rank one 
matrices $A = aa\T$ with $N(0,1)$-distributed entries of $a$. In addition to these four types,
we also use three types of scaled matrices with theoretical results on their restricted isometry
constants. These include matrices with 
$N(0, 1/m)$-distributed entries, with entries chosen uniformly in $\pm 1/\sqrt{m}$, and finally with distribution
\begin{equation}\nonumber
  A_{ij} = \begin{cases}
    + \sqrt{3/m} & \text{with probability } \frac{1}{6}, \\
    0 & \text{with probability } \frac{2}{3}, \\
    - \sqrt{3/m} & \text{with probability } \frac{1}{6}.
  \end{cases}
\end{equation}
For each type we used \mbox{MATLAB 8.3.0} to randomly generate nine
matrices of three different sizes with the number of rows ranging from 15
to 30, number of columns
from 25 to 40, and order $k$ between three and five, see Table~\ref{matrixsizes}.

@@

The tests were performed on a Linux cluster with Intel i3 CPUs with 3.2GHz, 4MB cache, and 8GB memory running
Linux. Each computation was performed single-threaded with a single process running on each computer and with 
a time limit of four hours. The code was compiled with gcc 4.4.5 with \texttt{-O3} optimization. For 39 matrices, 
the evaluation of~\eqref{Asp08Dual} was aborted for all 15 choices of $\rho$, because of the memory limit of 8GB.
\color{red}We also failed to solve \eqref{MISDP} for the left-hand side of the three medium sized rank one matrices with default
parameters because SDPA failed to solve the root node relaxation, however with different settings or a different
SDP-solver we could verify that $\alpha_k$ equaled zero up to the numerical tolerance for these instances. \color{black}
These are counted with the maximum allowed time of four hours.

@@

The restricted isometry constants of all matrices can be found in the appendix. Note that for the unscaled matrices,
the restricted isometry constants can be expected to be much larger than for the scaled ones. Even for the scaled
matrices, the restricted isometry property did not hold for any of the matrices in the testset, since the instances
were too small to satisfy the needed relation between $n$ and $k$.
\color{red} Furthermore, since we solved
the instances with a feasibility tolerance of $10^{-6}$, which influences the optimal objective by a factor of the
largest objective coefficients and the number of variables, the computed constants are only exact up to a factor
of about $10^{-2}$ for the unscaled matrices, while the precision is much better for the normalized ones because of their
smaller objective coefficients.\color{black}

@@

\noindent
This paper shows that exact values of restricted isometry constants can be
computed via mixed integer semidefinite programming. The computed bounds
significantly improve the bounds computed by the semidefinite
relaxations~\eqref{Asp07} and~\eqref{Asp08}
proposed in~\cite{Asp07} and \cite{Asp08}, respectively. In conclusion, the
relaxation~\eqref{Asp07}
is fast to compute, but produces a gap of around 90\,\% for
$\alpha_k$ and 30\,\% for $\beta_k$. If successfully solved, 
\eqref{Asp08} provides better bounds for $\beta_k$ for most types of 
matrices, but its running time is too large
to be used in practice. Moreover, directly solving~\eqref{MISDP} is often
faster and produces the exact value. Consequently, 
\eqref{Asp08} should not be used in practice.

@@

\appendix{ \input{tables/matrixList}}

>>>>

\noindent
In this section we demonstrate the applicability of the proposed MISDP-formulation and use it to
assess the quality of the relaxations proposed in~\cite{Asp08} and~\cite{Asp07}.
For solving the mixed integer semidefinite programs, we use the MISDP solver SCIP-SDP~\cite{SCIP-SDP,GalPU16},
originally developed in~\cite{Mar13} and~\cite{MS12}. It combines the
branch and bound framework SCIP 3.2.1 \cite{SCIP} with 
interior-point SDP-solvers, in our case SDPA~7.3.8~\cite{SDPA6,SDPA7}.

@@

We use a testset consisting of 63 matrices of the following seven
different types: band matrices,
with band size three or five with entries within the band chosen uniformly in $\{0,1\}$, binary matrices with all entries chosen 
uniformly in $\{0,1\}$, Gaussian-distributed matrices, and rank one 
matrices $A = aa\T$ with $N(0,1)$-distributed entries of $a$. In addition to these four types,
we also use three types of matrices with scaled parameters. These include matrices with 
$N(0, 1/m)$-distributed entries, with entries chosen uniformly in $\pm 1/\sqrt{m}$, and finally with distribution
\begin{equation}\nonumber
  A_{ij} = \begin{cases}
    + \sqrt{3/m} & \text{with probability } \frac{1}{6}, \\
    0 & \text{with probability } \frac{2}{3}, \\
    - \sqrt{3/m} & \text{with probability } \frac{1}{6}.
  \end{cases}
\end{equation}
For each type we used \mbox{MATLAB 8.3.0} to randomly generate nine
matrices of three different sizes with the number of rows ranging from 15
to 30, number of columns
from 25 to 40, and order $k$ between three and five, see Table~\ref{matrixsizes}.

@@

The tests were performed on a Linux cluster with Intel i3 CPUs with 3.2GHz, 4MB cache, and 8GB memory running
Linux. Each computation was performed single-threaded with a single process running on each computer and with 
a time limit of four hours. The code was compiled with gcc 4.4.5 with \texttt{-O3} optimization. For 39 matrices, 
the evaluation of~\eqref{Asp08Dual} was aborted for all 15 choices of $\rho$, because of the memory limit of 8GB.
We also failed to solve \eqref{MISDP} for the left-hand side of the three medium sized rank one matrices with default
parameters, because SDPA failed to solve the root node relaxation. With different settings or a different
SDP-solver, however, we could verify that $\alpha_k$ is zero for these instances.
These are counted with the maximum allowed time of four hours.

@@

The restricted isometry constants of all matrices can be found in the
appendix. Note that for the unscaled matrices (the first 36 instances until
instance \texttt{rnk140403C}),
we expect the restricted isometry constants to be much larger than for the scaled ones. Even for the scaled
matrices, the restricted isometry property never holds for the matrices in the testset, since the instances
are too small to satisfy the needed relation between $n$ and $k$.

@@

\noindent
This paper shows that exact values of restricted isometry constants can be
computed via mixed integer semidefinite programming. The computed bounds
significantly improve the bounds computed by the semidefinite
relaxations~\eqref{Asp07} and~\eqref{Asp08}
proposed in~\cite{Asp07} and \cite{Asp08}, respectively. In conclusion, the
relaxation~\eqref{Asp07}
is fast to compute, but produces a gap of around 90\,\% for
$\alpha_k$ and 30\,\% for $\beta_k$. If successfully solved, 
\eqref{Asp08} provides better bounds for $\beta_k$ for most types of 
matrices, but its running time is too large
to be used in practice. Moreover, directly solving~\eqref{MISDP} is often
faster and produces the exact value!

@@

\newpage
\appendix{ \input{tables/matrixList}}
